Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S.

### 2023-04-23 Draft.Rmd
popop

### Models 
### 1. Linear Regression
For linear regression (LR), the unscaled training set was used since variance does not affect the prediction performance of this model. Step Akaike Information Criteria (AIC) was used to simplify feature amount without impacting the model's performance. Features are dropped as the AIC score decreases and feature dropping stops once the AIC score increases. 'CMAQ'and 'aod' were determined to be the most important features, and these predictors were used in the recipe. Then, model was fitted. Using the summary function, we found the LR model had a RMSE of 2.27 performance. After 10-fold cross validation, the RMSE score was over 2, indicating that this model's prediction performance is not good.

@@

Based on the test set performance, it seems that our model gives the closest prediction for both Iowa and Virginia  The model gives the furthest prediction for Oregon. We hypothesize that the model gives bad performance for Oregon because despite being clearly larger than both Iowa and Virginia, it has the same number of monitors as the latter states, making the data less representative of the variation in PM2.5 concentration for the state.

@@

Indiana, Ohio, Iowa, and Alabama most frequently compared to the other states. Our model also gives the furthers predictions for California, Colorado, Kentucky, Pennsylvania, and Texas most frequently. We noticed that some of these states where our model's prediction performance was bad were on the larger side. We hypothesize that the reason why our prediction performance is bad could be attributed to the possible increase of variance in ambient air pollution that comes with a larger state area or that the model has poor outlier handling.

>>>>

### Models 
### 1. Linear Regression
For linear regression (LR), the unscaled training set was used since variance does not affect the prediction performance of this model. Step Akaike Information Criteria (AIC) was used to simplify feature amount without impacting the model's performance. Features are dropped as the AIC score decreases and feature dropping stops once the AIC score increases. 'CMAQ'and 'aod' were determined to be the most important features, and these predictors were used in the recipe. Then, model was fitted. Using the summary function, we found the LR model had a RMSE of 2.27 performance. After 10-fold cross validation, the RMSE score was over 1, indicating that this model's prediction performance is not good.

@@

Based on the test set performance, it seems that our model gives the closest prediction for both Iowa and Virginia  The model gives the furthest prediction for Oregon. We hypothesize that the model gives bad performance for Oregon because despite being clearly larger than both Iowa and Virginia, it has the same number of monitors as the latter states, making the data less representative of the variation in PM2.5 concentration for the state for the model to make a good prediction based on the features we chose.

@@

'CMAQ', 'county_area', and 'aod' might predict where our model performs well or not. We define a good prediction by having an absolute difference of less than 0.5 between 'value' and '.pred'. Indiana, Ohio, Iowa, and Alabama most frequently have good predictions. So the model does fairly well at predicting some of the Midwest. Our model gives the furthest predictions for California, Colorado, Kentucky, Pennsylvania, and Texas most frequently. So our model does poorly at predicting the West and the South. We think that including the variables 'popdens_county', 'zcta', and 'log_nei_2008_pm25_sum_10000' might improve the model performance since they were the following variables listed as most important in our feature selection process.
